Module analysis

### 2023-01-31 report/cs_analysis.Rmd
Add varying Z for various interaction strengths

>>>>

# SEM

```{r}
library(piecewiseSEM)
library(semEff)
```

```{r}
df_sem <- df %>%
  filter(stab_com < 1000) %>%
  mutate(across(c(sync, avg_cv_sp, stab_com, richness, avg_max_int, gini_max_int), log, , .names = "log_{.col}"))
```

```{r}
pairs(
  df_sem %>%
    select(ct, log_richness, max_tlvl,
      log_gini_max_int, log_avg_max_int, log_sync,
      log_avg_cv_sp, log_stab_com),
    pch = 19)
```

```{r}
sem_list <- list(
  max_tlvl = lm(max_tlvl ~ log_richness + ct, df_sem),
  gini_max_int = lm(gini_max_int ~ log_richness + ct + max_tlvl, df_sem),
  log_avg_max_int = lm(log_avg_max_int ~ log_richness + ct + max_tlvl, df_sem),
  log_sync = lm(log_sync ~ ct + log_richness + max_tlvl + gini_max_int, df_sem),
  log_avg_cv_sp = lm(log_avg_cv_sp ~ ct + log_richness + max_tlvl +  gini_max_int, df_sem),
  log_stab_com = lm(log_stab_com ~ log_sync + log_avg_cv_sp, df_sem)#~,
  #log_avg_max_int %~~% log_richness
)

sem <- as.psem(sem_list)
```

```{r}
map_dfr(setNames(sem_list, names(sem_list)), performance::check_collinearity)
map_dfr(sem_list, rsquared)
summary(sem)
```

```{r}
plot(sem)
```

```{r}
boot_eff <- semEff(sem, R = 50, ci.type =  "perc")
summary(boot_eff)
```

```{r}
sem_int_list <- list(
  max_tlvl = lm(max_tlvl ~ log_richness + ct, df_sem),
  gini_max_int = lm(gini_max_int ~ log_richness + ct + max_tlvl, df_sem),
  log_avg_max_int = lm(log_avg_max_int ~ log_richness + ct + max_tlvl, df_sem),
  log_sync = lm(log_sync ~ log_richness + log_avg_max_int + gini_max_int, df_sem),
  log_avg_cv_sp = lm(log_avg_cv_sp ~ log_richness + log_avg_max_int + gini_max_int, df_sem),
  log_stab_com = lm(log_stab_com ~ log_sync + log_avg_cv_sp, df_sem)
)

sem_int <- as.psem(sem_int_list)
plot(sem_int)
```

```{r}
map(setNames(sem_int_list, names(sem_int_list)), performance::check_collinearity)
map_dfr(sem_list, rsquared)
summary(sem)
```

```{r}
boot_int <- semEff(sem_int, R = 50, ci.type =  "perc")
summary(boot_int)
```

### 2023-01-31 report/module_analysis.Rmd
Add varying Z for various interaction strengths

---
title: big stoch sim
author: alain
date: 2022-09-12
format:
  html:
    embed-resources: false
---

```{r}
library(arrow)
library(tidyverse)
library(magrittr)
```

```{r}
pq_brut <- open_dataset(here::here("res", "sim_module.arrow"), format = "arrow")
df <- pq_brut %>%
  collect()
colnames(df)
```

```{r}
df %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module, y = values, color = as.factor(alphaij))) +
  geom_boxplot()+
  facet_grid(
    rows = vars(stab),
    scales = "free_y")
```

>>>>

---
title: Module analysis
author: alain
date: 2022-09-12
output:
  bookdown::pdf_document2
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")

theme_set(mytheme)
```

# Stochastic natural death

## Competition

```{r}
comp_brut <- open_dataset(here::here("res", "sim_module_competition.arrow"), format = "arrow")
comp <- comp_brut %>%
  collect() %>%
  select(where(is.double))
colnames(comp)
```

```{r, eval = FALSE}
unique(comp$env_stoch)
summary(comp$total_biomass)
```

```{r}
comp %>%
  filter(env_stoch == 0.9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = as.factor(richness), y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(
    rows = vars(stab), cols = vars(rho),
    scales = "free_y") +
  labs(x = "Species richness", y = "Stability component", color = "Interspecific competition")
```

- Main results:
  - Species richness increases stability
  - Interspecific competition does not affect stability expect when
    interspecific competition equals to intraspecific competition (decrease the effect of species richness).
  - Interspecific competition desynchronises primary producers but increases
    varibility
  - The decrease of response diversity decreases the effect of species richness on stability
  - The decrease of response diversity increase synchrony but decrease
    population variability

## Modules

```{r}
mod_brut <- open_dataset(here::here("res", "sim_module_d.arrow"), format = "arrow")
mod <- mod_brut %>%
  collect() %>%
  select(where(is.numeric) | starts_with("module")) %>%
  filter(richness / S == 1.0)
colnames(mod)
```

```{r, eval = FALSE}
unique(mod$env_stoch)
summary(as_tibble(mod))
```

```{r}
mod %>%
  filter(rho == 0, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")
```

```{r}
mod %>%
  filter(rho == 0.5, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")

```

```{r}
mod %>%
  filter(rho == 1, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")
```
